A boundary integral equation approach to computing eigenvalues of the Stokes operator

### 2019-12-10 doc/review-acom/response-to-referees.tex

We have also marked in {\color{red} red} a couple of
changes around the explanation of the order of convergence
of the generalized Gaussian quadrature rule we used.
The rule is designed to be 16th order, i.e. it is designed
to integrate $p(y)\log|x-y| + q(y)$ for $p$ and
$q$ polynomials of degree up to 15. Our kernel is actually
smoother than $\log|x-y|$ (with leading order
singularity $|x-y|^2\log|x-y|$) and in our experience the
convergence rate is typically higher (20th order in
the example in the paper). In the revised text, we have clarified
that the rule is not designed to be 20th order but that this is
the behavior we observed.

\section{Reviewer 1}

A general concern of the first reviewer was the discussion
of what we originally called ``spurious double roots''.
We appreciate the reviewer pressing us on this issue, as it
has clarified the results of the paper.

To explain what happened originally, we computed the
roots using a rootfinding
utility in another software package, \texttt{chebfun}. In our
numerical experiments, we were puzzled that the output of the
utility consistently gave us pairs of roots which seemed to
be identical. Yet, they did not actually seem to correspond to
a double root of the Fredholm determinant in any meaningful sense
(indeed, the derivative of our interpolant was not particularly
close to zero for these points). We performed some other tests
(including checking whether there truly was a corresponding
two-dimensional eigenspace) and no other tests suggested a double
root existed. Thus, in the original submission, we made a note
suggesting that rounding errors may explain this strange behavior
and we described our ad hoc methods for determining true roots.
To confuse matters, we mentioned the existence of roots that
were computed and had large imaginary part as being part of
the same phenomenon (they are not).

@@

We've included a sketch of the proof below.

The main estimate needed is to show that the error in computing
the Fredholm determinant at an eigenvalue of the Stokes operator
is directly proportional to the quadrature error in 
evaluating the integral operator applied to the null vector (A proof of this result for the 
Laplace case is contained in~\cref{sec:appb}). 
Based on the second-kind nature of the operator from $C^{m,\alpha} \to C^{m,\alpha}$
for $m<\ell$ and $\alpha>0$ (when the boundary is $C^{\ell}$) one can show
that the resulting null vector must be at least in $C^{m,\alpha}$.
The quadrature error using Kress quadratures converges algebraically
when both the boundary and density being approximated are $C^{m,\alpha}$.

\item[4.] {\color{blue} In the paper it is not explicitly mentioned that
  both presented boundary integral equations may have non-trivial null-spaces
  for certain $k$ with ${\rm Im}(k) < 0$. These $k$s correspond to the eigenvalues of some
  related eigenvalue problems in the exterior domain. It is possible that
  these $k$s are also be detected by the root search method and that these
  eigenvalues may be some of those “spurious eigenvalues too far from
  real-valued” which are blamed to be caused by “noisy numerical determinant
  evaluation” but are actually caused by the used boundary integral formulation.}

Thank you for pointing this out. This is indeed the case, as can be seen
by the dependence of many of the results on ${\rm Im}(k) \geq 0$. We have
added a remark about this after Theorem 10 and updated the theorems and lemmas
as suggested in other comments below. These are indeed detected by the
root finding procedure, though they are easy to identify as
being distinct from the real-valued roots which correspond to
eigenvalues (the imaginary part of the real-valued roots is several
orders of magnitude smaller than the imaginary part of the
spurious roots we observed).

@@

\item[5.] {\color{blue}
  In general it would be of interest which kind of spurious eigenvalues may
occur from a“noisy numerical determinant evaluation”. In the paper
only two kinds are specified, namely spurious eigenvalues with “large”
imaginary part (which may result also from the boundary integral for-
mulation itself as pointed out above) and spurious “double roots” on
the real line. Isn’t it possible that for example spurious single roots
occur on the real line?}

Again, thank you for pressing this point. We believe that this
is mostly resolved by the discussion above. In the case of a spurious
single root on the real line, we imagine that this could happen, though it
was not observed, i.e. one can imagine a pathological domain for which a
spurious root exists that is within, say, $10^{-7}$ of the real line.
Such a root would be hard to distinguish from a true root.
In all cases, we also computed the smallest singular
value of the BIE corresponding to the real part of a given root.
This singular value was found to be quite small. Because we believe
that the integral equation was well-resolved
by our discretization, this is rather convincing evidence that an
eigenvalue of the PDE has been found.

### 2019-12-10 doc/draft-01-stokes/appendix-b.tex
minor changes
adding small chnges
finishing appendix b

\section{Proof of convergence for general case}
Let $T_{n}$ denote the space of trignometric polynomials
$\{ e^{-inx}, \ldots e^{inx} \}$. 
Let $\mathcal{K}$ denote the Helmholtz double layer integral operator
given by
\begin{equation}
\mathcal{K}[\phi] = \frac{i}{2}\int_{\Gamma} k H_{1}^{1}(k|x-y|) \frac{(x-y)\cdot n(y)}{|x-y|} \phi(y) dS_{y} \,
\end{equation}
where $k$ the Helmholtz parameter is assumed to be a dirichlet eigenvalue of the simply connected
domain $\Omega$ whose boundary is given by $\Gamma$, and $H_{1}^{1}$ is the Hankel function
of the first kind of order $1$.
It is well known that if $\Gamma$ is of class $C^{\ell,\alpha}$, then the
integral operator $\mathcal{K}: C^{\ell-1,\alpha} \to C^{\ell-1,\alpha}$ is a compact operator.
Furthermore, as shown in Zhao and Barnett, at the simple Dirichlet eigenvalues of the operator
$I + \mathcal{K}$ is not invertible and has a one dimensional null-space. Let $\phi_{0}$ denote
the null vector. 
By Fredholm theory, the adjoint $I + \mathcal{K}^{\ast}$ also has a one-dimensional null space.
Let $\psi_{0}$ denote the null vector of $I+\mathcal{K}^{\ast}$.

@@

And finally the operator $C_{n}$ is the restriction of $I_{n} + K_{n}$ as a map from
$P_{n} \phi_{0} \to P_{n} \psi_{0}^{\perp}$ and is its adjoint is given by
\begin{equation}
C_{n}^{\ast}\psi = \left( (I_{n} + K_{n}^{\ast}) \left(P_{n} \psi - (P_{n} \psi, P_{n} \psi_{0}) P_{n} \psi_{0} \right), P_{n} \phi_{0} \right)
\end{equation}

In this setting, $\phi_{0}$ and $\psi_{0}$ are known to be linearly independent.
Thus for large enough $n$, 
$f_{n} = P_{n} \phi_{0} - (P_{n} \phi_{0}, P_{n} \psi_{0}) P_{n} \psi_{0} \neq 0$
and $f_{n} \in P_{n} \psi_{0}^{\perp}$.
Similarly $g_{n} = P_{n} \psi_{0} - (P_{n}\phi_{0}, P_{n} \psi_{0}) P_{n} \phi_{0} \neq 0$
and $g_{n} \in P_{n} \phi_{0}^{\perp}$.

By block linear algebra, the determinant of the matrix above is also given by
\begin{equation}
\det{
\begin{bmatrix}
A_{n} & B_{n} \\
C_{n} & D_{n} 
\end{bmatrix} 
}
= \det{D_{n}} \cdot (A_{n} - C_{n} D_{n}^{-1} B_{n}) \,.
\end{equation}
Note that $C_{n} D_{n}^{-1} B_{n}$ is a number and can be evaluated as 
\begin{equation}
\begin{aligned}
C_{n} D_{n}^{-1} B_{n} &= 
\left( f_{n} \cdot  C_{n} D_{n}^{-1} B_{n} g_{n} \right) \\
&= \left( C_{n}^{\ast} f_{n}, D_{n}^{-1} B_{n} g_{n} \right) \\
&= \left( \left((I_{n} + K_{n}^{\ast})f_{n}, P_{n} \phi_{0}\right) P_{n} \phi_{0}, D_{n}^{-1} B_{n} g_{n} \right) \\
&= \left(f_{n}, (I_{n} + K_{n})P_{n} \phi_{0} \right) \cdot (P_{n} \phi_{0}, D_{n}^{-1} B_{n} g_{n}) \\
&= \left(f_{n}, (I_{n} + K_{n})P_{n} \phi_{0} \right) \cdot (P_{n} \phi_{0}, D_{n}^{-1} P_{n} \psi_{0})  \cdot 
\left((I_{n} + K_{n}^{\ast})P_{n} \psi_{0}, g_{n}  \right) \\
\end{aligned}
\end{equation}

First, we observe that $D_{n}$ is the restriction of an operator in trace class
with a kernel whose leading order singularity is $r^{2} \log{r}$. 
Thus, $D_{n}$ is itself in trace class, and extending the proof of Bornemann to this case,
we observe that the discrete determinant of $D_{n}$ converges to the Fredholm determinant
of $D$ at a rate of at least $O(1/n)$. This in particular implies that
$\sup_{n} |\det{D_{n}}| <\infty$.
Using Cauchy-Schwarz, the quantities $A_{n}, C_{n}D_{n}^{-1}B_{n}$ are bounded by
$|(I_{n} + K_{n})\phi_{n}|$.
Combining these results together, we get that
\begin{equation}
\left| \det{\begin{bmatrix} A_{n} & B_{n} \\ C_{n} & D_{n} \end{bmatrix}} \right| \leq C |(I_{n} + K_{n})\phi_{n}|
\end{equation}

>>>>

\section{Sketch of proof}
\label{sec:appb}
Let $T_{n}$ denote the space of trigonometric polynomials
$\{ e^{-int}, \ldots e^{int} \}$. 
Let $\mathcal{K}$ denote the Helmholtz double layer integral operator
given by
\begin{equation}
\mathcal{K}[\phi] = \frac{i}{2}\int_{\Gamma} k H_{1}^{1}(k|x(t)-y(t)|) \frac{(x(t)-y(t))\cdot n(y(t))}{|x(t)-y(t)|} \phi(y(t)) dS_{y} \,
\end{equation}
where $k$ the Helmholtz parameter is assumed to be a dirichlet eigenvalue of the simply connected
domain $\Omega$ whose boundary is given by $\Gamma$, $(x(t),y(t)): [0,2\pi] \to \Gamma$ is
an arclength parameterization of the boundary, and $H_{1}^{1}$ is the Hankel function
of the first kind of order $1$.
It is well known that if $\Gamma$ is of class $C^{\ell,\alpha}$, then the
integral operator $\mathcal{K}: C^{\ell-1,\alpha} \to C^{\ell-1,\alpha}$ is a compact operator.
Furthermore, as shown in Zhao and Barnett, at the simple Dirichlet eigenvalues of the operator
$I + \mathcal{K}$ is not invertible and has a one dimensional null-space. Let $\phi_{0}$ denote
the null vector. 
By Fredholm theory, the adjoint $I + \mathcal{K}^{\ast}$ also has a one-dimensional null space.
Let $\psi_{0}$ denote the null vector of $I+\mathcal{K}^{\ast}$.

In order to prove the regularity of the eigenfunctions, we need to analyze
the kernels in $C^{\ell-1,\alpha}$ space. 
However in order to obtain a bound for the magnitude of the Fredholm
determinant, it is more convenient to work in $L^{2}$. It is well known
that the integral operator $I + \mathcal{K}$ is a second kind
Fredholm operator from $L^{2} \to L^{2}$ as well.
For the remainder of the section, we will work with $L^{2}$.

@@

And finally the operator $C_{n}$ is the restriction of $I_{n} + K_{n}$ as a map from
$P_{n} \phi_{0} \to P_{n} \psi_{0}^{\perp}$ and is given by
\begin{equation}
C_{n}P_{n} \phi_{0} = (I_{n} + K_{n}) P_{n} \phi_{0} - \left( (I_{n} + K_{n}) \phi_{0}, P_{n} \psi_{0} \right)P_{n} \psi_{0}
\end{equation}

%Thus for large enough $n$, 
%$f_{n} = P_{n} \phi_{0} - (P_{n} \phi_{0}, P_{n} \psi_{0}) P_{n} \psi_{0} \neq 0$
%and $f_{n} \in P_{n} \psi_{0}^{\perp}$.
%Similarly $g_{n} = P_{n} \psi_{0} - (P_{n}\phi_{0}, P_{n} \psi_{0}) P_{n} \phi_{0} \neq 0$
%and $g_{n} \in P_{n} \phi_{0}^{\perp}$.

By block linear algebra, the determinant of the matrix above is also given by
\begin{equation}
\det{
\begin{bmatrix}
A_{n} & B_{n} \\
C_{n} & D_{n} 
\end{bmatrix} 
}
= \det{D_{n}} \cdot (A_{n} - B_{n} D_{n}^{-1} C_{n}) \,.
\end{equation}

First, we observe that $D_{n}$ is the restriction of an operator in trace class
with a kernel whose leading order singularity is $r^{2} \log{r}$. 
{\color{red} Thus, $D_{n}$ is itself in trace class, and extending the proof of Bornemann to this case,
we observe that the discrete determinant of $D_{n}$ converges to the Fredholm determinant
of $D$ at a rate of at least $O(1/n)$. This in particular implies that
$\sup_{n} |\det{D_{n}}| <\infty$}.
Using Cauchy-Schwarz, we get that
\begin{equation}
\begin{aligned}
|A_{n}| &\leq |P_{n} \psi_{0}|_{2} |(I_{n} + K_{n})P_{n} \phi_{0}|_{2} \\
|B_{n}D_{n}^{-1} C_{n}| &\leq |B_{n}||D_{n}^{-1}|(1+ |P_{n} \psi_{0}|^2) |(I_{n} + K_{n})P_{n} \phi_{0}| \,.
\end{aligned}
\end{equation}
$B_{n}$ is a discretized block of a bounded operator and hence is bounded, 
and $D_{n}^{-1}$ is a bounded operator by the discussion above.

Combining these results together, we get that
\begin{equation}
\left| \det{\begin{bmatrix} A_{n} & B_{n} \\ C_{n} & D_{n} \end{bmatrix}} \right| \leq C |(I_{n} + K_{n})\phi_{n}| \, ,
\end{equation}
where $C = \det{D_{n}} |B_{n}||D_{n}^{-1}| (1 + |P_{n} \psi_{0}|^2)$.
